Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-01-23 2017-Xudong-SMT/FM-2018-extended.tex
partial polish of appendix

\paragraph*{Algebra presentations.}
In practice, the parameterization of the pNet model by some specific
action algebra is realized by the definition of an ``algebra
presentation''. It will be used to check the
well-formedness of a pNet system, and to define the translation of the pNet
semantics into the SMT engine input language.

@@

We check the satisfiability of each open transitions
using the SMT solver Z3. More precisely, we translate
the OT predicate into an SMTlib assertion in the context provided by
the possible state variable assignments in the source state. In this
section, we describe the translation of the algebra presentation, of
assignments, and of the predicates.

Our implementation submits satisfiability requests to
Z3 using its JAVA API. Here, for readability, we
show the Z3 code using its SMT-LIB input language.
Note that in the previous sections, the OTs were displayed in a
  simplied, human readable form. The input and output of our tool, and
  also the generated SMTlib fragments, are
  slightly more difficult to read, in 
particular use structured names for the fresh variables generated
by the algorithm, allowing tracability of the result (see
appendix~\ref{appendix:variables}).

@@

To build the input submitted to Z3 for each OT,
we translate the algebra presentation, the predicates and the
variable assignments (the Post part) into Z3 (Java-API) syntax.
% The $presentation$ and the predicates are translated separately.

\paragraph{Translation of action algebra presentation.}
In App.~\ref{appendix:mapsmt}, we formally define the condition of
well-formedness of an algebra presentation, and its translation into
SMTlib declarations (\texttt{declare-datatypes} and
\texttt{declare-fun}). It ensures that the generated code is correct,
and will not raise runtime errors in the SMT engine. Remark that the
\texttt{declare-datatypes} command comprises both the action
constructors from table \ref{table:BIPalgebra} and also the constant action
names from our example.

@@

\paragraph{Checking satisfiability.}

In App.~\ref{appendix:mapsmt} we formally define all steps of the translation of each
open transition, including:
%
\begin{itemize}
\item collect all variables in the transition, and declare them,
\item check well-formedness and correct typing of expressions,
\item translate the predicate into a conjunction of assertions,
\item if present in the source state, translate the state-variable
  assignments into a disjunctive assertion.
\end{itemize}
%
Here again, this translation ensures that no runtime error will occur
when running the SMT engine.

Figure~\ref{schema:smt-lib} shows the decomposition of the
predicate into a set of asserts, each encoding an elementary equality,
inequality, or a guard. 
The result (sat or unsat) of the final command in the translation (\texttt{check-sat}) is decoded by
our algorithm.

@@

\begin{enumerate}
\item[A] pNet semantic rules
\item[B] Variable management
\item[C] Algebra presentations
\item[D] Well-formness and typing of expressions
\item[E] Translation to SMTlib input language
\item[F] Full output for our use-case

@@

\section{Variable management}
\label{appendix:variables}
The variables in each synchronisation vector are considered local: for
a given pNet expression, we must have fresh local variables for each
occurrence of a vector (= each time we instantiate rule
Tr2). Similarly the state variables of each copy of a given pLTS in
the system, must be distinct, and those created for each application
of Tr2 have to be fresh and all distinct. This will be implemented
within the open-automaton generation algorithm, e.g. using name
generation using a global counter as a suffix.

\paragraph{Fresh variables}
Application of the semantic rules require generating a lot of fresh
names, for different kind of variables. We could use a global name
generator to guarantee unicity, but 
at the same time, we also hope them still
readable. Here we rename the variables with a regular format using the fresh function.
More precisely, the fresh function generates a new name adding a
suffix after the original name. The suffix contains three parts
combined by a colon.

@@

@@

		&       		\\\specialrule{0em}{5pt}{5pt}
		(Overloaded FUN)							
		&					
		& 			\\\specialrule{0em}{3pt}{3pt}
		$\dfrac{\EPres \vdash \texttt{FUN} :: A_1,...,A_n \rightarrow Action ~~\EPres \vdash A_1~~...~~\EPres \vdash A_n ~~\EEnv \vdash x_1:A_1~~...~~\EEnv \vdash x_n:A_n}{\EPres,\EEnv \vdash \texttt{FUN}(x_1,...,x_n):Action}$ 
		& 
		&			\\
		\specialrule{0em}{5pt}{5pt}\hline
	\end{tabular}
\end{table}

@@

\section{Management of state variables}
\label{appendix:translating-assignments}
In a pLTS, there may be several incoming transitions 
that assign potentially different values to a state variable.
To handle such cases, the algorithm manages for each pLTS state, as a
list of expressions collected from the assignments of each state
variable $v$ (as registered in the incoming OTs).
%% The list we used contains triples <$v$, $S$, $AssignRH$> 
%% where $v$ is the variable in pLTS, $S$ is its owner state and
%% $AssignRH$ is a list of expressions over other variables in the right
%% hand side of assignments of $v$.
% As a pragmatic extension to the formal definition, we also manage
% ``global variables'', defined in all states of the pLTS.
For a global state in the open automaton, getting the set of state
variables (which may be used in a transition) will simply be the union
of state variables of the individual pLTS states constituting this
global state, as variables of different pLTS are distinct.

\section{Translation to SMTlib input language}
\label{appendix:mapsmt}
The pNet elements, as defined above, can be full translated into
SMT-LIB language (\ref{SMTlibv2.5}), but there are a number of differences in the
structure of the models/languages, so the translation is not trivial.

In this section we shall define separately the translation of the
extended algebra presentation for one pNet (so it will be common for
the study of all OTs of one use-case); and the translation of each
predicate (of an open-transition) of this pNet.

@@

\vspace{1ex}\noindent
%\begin{table}\caption{\leftline{Mapping}}
	\begin{tabular}{p{3cm}p{9cm}}
		\hline\specialrule{0em}{1pt}{1pt}			
		(Presentation)							
		&								\\\specialrule{0em}{1pt}{1pt}		
		&$Sorts \& Constructors \translate$	declare-datatypes				\\\specialrule{0em}{1pt}{1pt}
		&$Operators \lhook\joinrel\longrightarrow$	declare-function		\\\specialrule{0em}{1pt}{1pt}
		(Predicate)							
		&								\\\specialrule{0em}{1pt}{1pt}
		&$\EEnv ~~\lhook\joinrel\longrightarrow$	declare-const		\\\specialrule{0em}{1pt}{1pt}
		&$Pred ~~\lhook\joinrel\longrightarrow$	 assert		\\\specialrule{0em}{1pt}{1pt}
		&$\texttt{FUN}(x_1,...,x_n)$		\\\specialrule{0em}{1pt}{1pt}
		&$\texttt{pNet expression} ~~\lhook\joinrel\longrightarrow$ SMTLib expression	\\\specialrule{0em}{1pt}{1pt}
		&$...$	\\\specialrule{0em}{1pt}{1pt}
		\specialrule{0em}{1pt}{1pt}\hline
	\end{tabular}
%\end{table}

@@

In the following definitions, we use abstract functions corresponding to
SMTlib/Z3 constructs. In practice they can be implemented either as
SMTLib scripting programs, or as calls to the Z3 java API.

\subsubsection{Presentation Translation} \TODO{Duplicate ?}
We define here the translation of the algebra presentation, extended
with the constant operators collected from the pNet. It produces the
declaration of sorts (excepted Bool and Int) with their constructors
and selectors as \texttt{declare-datatypes}, and the declaration of
auxiliary operators as \texttt{declare-function} constructs.
For sorts, we must distinguish the case of mutually defined sorts
(e.g. edges and vertices in a graph), that must be declared within a
single \texttt{declare-datatypes} construct. For this we define:

\begin{definition}
Define the (strict) order "is-using" between sorts $S1 is-using S2$ iff $S2$
occurs as the sort of one argument in the constructors of $S1$.
\end{definition}

\begin{lstlisting}
Let Pres = <Sorts, Constrs, Ops> an extended presentation
(i.e. including the constants from the pnet),
- define MySorts = Sorts \ {Bool, Int}
- compute the strongly connected components in the graph
  of MySorts with respect to the relation is-using
- for each SCC in this graph, construct: 
  datatype-declaration = TrPresentation (Pres, SCC)
- for each other operator op in Ops, construct:
  function-declaration = TrPresentation (Pres, op)
\end{lstlisting}

@@

		& 			\\\specialrule{0em}{1pt}{1pt}
		$\dfrac{n = arity(constr) \neq 0 \qquad |sels|=n
                  \qquad sels =  BuildSels(constr)}
                       {constr \translate \texttt{(name($constr$) . sels)}}$

@@

		& 			\\\specialrule{0em}{1pt}{1pt}
		$\dfrac{op : sort_1, ..., sort_n : sort}
                       {op \translate \texttt{(declare-fun name($op$) ($sortname_1
                           ... sortname_n$) $sortname$)}}$

@@

		& 			\\\specialrule{0em}{1pt}{1pt}
		$\dfrac{}
                       {FUN \translate \texttt{map $\translate CollectFunTypes(pNet)$ }}$

		&&       		\\\specialrule{0em}{1pt}{1pt}
		$\dfrac{}
                       {argstypes \translate \texttt{(declare-fun
                           $BuildFunInstance(argstypes)$ (map name
                           $argstypes$) Action) }}$

@@

\smallskip
Where:
\begin{itemize}
  \item the $BuildSels$ function, for a constructor with $n$ arguments,
argument sorts $sort_i$, and selector names $sel_i$, builds the list
$\{(sel_i, sort_i)\}_{i\in[1..n]}$.
\item the $CollectFunTypes$ function collects all possible instances
  of the overloaded FUN argument types found in the pNet, as computed by the typing
  rules; and $BuildFunInstance$ use these argument types as suffixes
  to disambiguate 
\end{itemize}

\begin{example}
  For the CCS presentation above, we would get (in SMTLib syntax):
\begin{lstlisting}
  (declare-datatypes Data ())
  (declare-datatypes Channel ())
  (declare-datatypes ()
      ((Action (Emit (Chan_E Channel) (Value_E Data))
               (Receive (Chan_R Channel) (Value_R Data))
               Tau )))
  \end{lstlisting}
\end{example}

\TODO{Need to add an example with 2 instances of FUN}

@@

\subsection{properties}
The rules in the sections above are meant to guarantee that no runtime
errors would occur when running Z3 from our algorithm. More precisely,
we ensure that:
\begin{itemize}
\item all datatypes, functions, variable used are declared,
\item datatypes declared are well-founded,
\item all expressions in declarations and assertions are well-typed.
\end{itemize}

\section{Full output for the FailureTimer open automaton}
\label{appendix:fullOutput}

\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, frame=single]
	{s0---failure(true)-->s1, t0---start(true)-->t1},{---:hb_B:13:1-->},   
	(failure(true)=failure(b1:sva_SV0:1:1))
	/\(start(true)=start(b2:sva_SV0:1:1))
	/\(:hb_B:13:1=fail(b0:sva_SV0:1:1))/\((b1:sva_SV0:1:1=b2:sva_SV0:1:1)
	/\(~(b1:sva_SV0:1:1\/b2:sva_SV0:1:1)\/b0:sva_SV0:1:1))
	/\(_fail_=:ra:1:1),{t := Max}
OT ---------------------------------------------------------------------------
		<s0_t0>-----:ra:1:1----><s1_t1>

@@

	{s1---resume(true)-->s0, t1---resume(true)-->t0},{---:hb_B:16:2-->},   
	(resume(true)=resume(b1:sva_SV1:1:2))
	/\(resume(true)=resume(b2:sva_SV1:1:2))
	/\(:hb_B:16:2=resume(b0:sva_SV1:1:2))/\((b1:sva_SV1:1:2=b2:sva_SV1:1:2)
	/\(~(b1:sva_SV1:1:2\/b2:sva_SV1:1:2)\/b0:sva_SV1:1:2))
	/\(_resume_=:ra:1:2),{}
OT ---------------------------------------------------------------------------
	<s1_t1>-----:ra:1:2----><s0_t0>

@@

\subsection{Scaling up: several subsystems on the same bus}
\TODO{Not yet !}
\begin{figure}
    \centerline{\includegraphics[width=\linewidth]{XFIG/ScaleUp}}
    \caption{pNet encoding of the composition of two Failure Monitoring modules}  \label{schema:scaleup}
\end{figure}

\newpage
\begin{figure}
$\begin{array}{r@{}l}
    ot_1  = &\openrule{\{s0 \xrightarrow{\OTvar{fail(true)}} s1, t0 \xrightarrow{\OTvar{start(true)}} t1\}, ~~\{\xrightarrow{hb_1}\},~~\OTvar{fail(false)}=\OTvar{fail(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{start(true)}=\OTvar{start(}\OTvar{b}_2\OTvar{)}\OTland hb_1=\OTvar{fail(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_1=\OTvar{\underline{fail}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{t := Max\}}
    	   {\ostate{s0,t0} \xrightarrow{v_1} \ostate{s1,t1}}\\
\end{array}$
  \smallskip

@@

\newpage
\end{document}

>>>>

\paragraph*{Algebra presentations.}
In practice, the parameterization of the pNet model by some specific
action algebra is realized by the definition of an ``algebra
presentation''. It will be used to check the
well-formedness of a pNet system, and to define the translation of the pNet
semantics into the SMT engine input language (\cite{BarFT-RR-17}).

@@

We check the satisfiability of each open transitions
using the SMT solver Z3.
%More precisely, we translate
%the OT predicate into an SMTlib assertion in the context provided by
%the possible state variable assignments in the source state.
In this
section, we describe the translation of the algebra presentation, of
assignments, and of the predicates.

Our implementation submits satisfiability requests to
Z3 using its JAVA API. Here, for readability, we
show the Z3 code using its SMT-LIB input language.
Note that in the previous sections, the OTs were displayed in a
  simplied, human readable form. The input and output of our tool, and
  also the generated SMTlib fragments, are
  slightly more difficult to read, in 
particular because of structured names for the fresh variables generated
by the algorithm, allowing tracability of the result (see
appendix~\ref{appendix:variables}).

@@

\paragraph{Production of the SMT-lib code}
  To build the input submitted to Z3 for each OT,
we translate the algebra presentation, the predicates and the
variable assignments into Z3 (Java-API) calls.
% The $presentation$ and the predicates are translated separately.

\emph{Translation of action algebra presentation.}
In Appendix~\ref{appendix:mapsmt}, we  define the condition of
well-formedness of an algebra presentation, and its translation into
SMTlib declarations (\texttt{declare-datatypes} and
\texttt{declare-fun}). It ensures that the generated code is correct,
and will not raise runtime errors in the SMT engine. Note that the
\texttt{declare-datatypes} command comprises both the action
constructors from table \ref{table:BIPalgebra} and also the constant action
names from our example.

@@

\emph{Translation of open transitions}

In App.~\ref{appendix:mapsmt} we formally define all steps of the translation of each
open transition, including:
%
\begin{itemize}
\item collect all variables in the transition, and declare them (using \texttt{declare-const})
\item check well-formedness and correct typing of expressions,
\item translate the predicate into a conjunction of assertions,
\item if present in the source state, translate the state-variable
  assignments into a disjunctive assertion.
\end{itemize}
%
Here again, this translation ensures that no runtime error will occur
when running the SMT engine.

Figure~\ref{schema:smt-lib} shows the decomposition of the
predicate into a set of asserts, each encoding an elementary equality,
inequality, or a guard. 
The result (sat or unsat) of the final command in the translation (\texttt{check-sat}) is decoded by
)our algorithm.

@@

\begin{enumerate}
\item[A] pNet semantic rules
\item[B] Variable management
  \subitem B.1 fresh variables
  \subitem B.2 managment of state variables
\item[C] Algebra presentations
\item[D] Well-formness and typing of expressions
\item[E] Translation to SMTlib input language
\item[F] Full output for the FailureTimer open automaton
\item[G] Scaling up

@@

\section{Variable management}
In this section we give more details on two facets of variable
management in the tool: the generation of variables names, and the
management of state variable assignments.

\subsection{Fresh variables}
\label{appendix:variables}
The variables in each synchronisation vector are considered local: for
a given pNet expression, we must have fresh local variables for each
occurrence of a vector (= each time we instantiate rule
Tr2). Similarly the state variables of each copy of a given pLTS in
the system, must be distinct, and those created for each application
of Tr2 have to be fresh and all distinct. This will be implemented
within the open-automaton generation algorithm, e.g. using name
generation using a global counter as a suffix.

Application of the semantic rules require generating a lot of fresh
names, for different kind of variables. We could use a global name
generator to guarantee unicity, but 
at the same time, we also hope them still
readable. Here we rename the variables with a regular format using the fresh function.
More precisely, the fresh function generates a new name adding a
suffix after the original name. The suffix contains three parts
combined by a colon.

@@

\subsection{Management of state variables}
\label{appendix:translating-assignments}
In a pLTS, there may be several incoming transitions 
that assign potentially different values to a state variable.
To handle such cases, the algorithm manages for each pLTS state, as a
list of expressions collected from the assignments of each state
variable $v$ (as registered in the incoming OTs).
%% The list we used contains triples <$v$, $S$, $AssignRH$> 
%% where $v$ is the variable in pLTS, $S$ is its owner state and
%% $AssignRH$ is a list of expressions over other variables in the right
%% hand side of assignments of $v$.
% As a pragmatic extension to the formal definition, we also manage
% ``global variables'', defined in all states of the pLTS.
For a global state in the open automaton, getting the set of state
variables (which may be used in a transition) will simply be the union
of state variables of the individual pLTS states constituting this
global state, as variables of different pLTS are distinct.

@@

		&       		\\\specialrule{0em}{5pt}{5pt}
		(Overloaded FUN operator)							
		&					
		& 			\\\specialrule{0em}{3pt}{3pt}
		$\dfrac{\EPres \vdash \texttt{FUN} :: A_1,...,A_n \rightarrow Action ~~\EPres \vdash A_1~~...~~\EPres \vdash A_n ~~\EEnv \vdash x_1:A_1~~...~~\EEnv \vdash x_n:A_n}{\EPres,\EEnv \vdash \texttt{FUN}(x_1,...,x_n):Action}$ 
		& 
		&			\\
		\specialrule{0em}{5pt}{5pt}\hline
	\end{tabular}
\end{table}

@@

\section{Translation to SMTlib input language}
\label{appendix:mapsmt}
The pNet elements, as defined above, can be full translated into
SMT-LIB language (\cite{BarFT-RR-17}), but there are a number of differences in the
structure of the models/languages, so the translation is not trivial.

In this section we define separately
\begin{itemize}
\item The translation of the
extended algebra presentation for one pNet (so it will be common for
the study of all OTs of one use-case);
\item and the translation of each open transition of the pNet,
  including its environment (list of variables), its predicate, and
  the encoding of the possible values of the state-variables of its
  source global state.
\end{itemize}
The following table summarizes the main elements of the translation.

@@

\vspace{1ex}\noindent
%\begin{table}\caption{\leftline{Mapping}}
	\begin{tabular}{p{3cm}p{9cm}}
		\hline\specialrule{0em}{1pt}{1pt}			
		(Presentation)							
		&								\\\specialrule{0em}{1pt}{1pt}		
		&$Sorts \& Constructors \translate$	declare-datatypes				\\\specialrule{0em}{1pt}{1pt}
		&$Other operators \lhook\joinrel\longrightarrow$	declare-function		\\\specialrule{0em}{1pt}{1pt}
		(Open Transition)							
		&								\\\specialrule{0em}{1pt}{1pt}
		&$\EEnv ~~\lhook\joinrel\longrightarrow$	declare-const		\\\specialrule{0em}{1pt}{1pt}
		&$Pred ~~\lhook\joinrel\longrightarrow$	 assert		\\\specialrule{0em}{1pt}{1pt}
		&$\texttt{FUN}(x_1,...,x_n)$	\TODO {Xudong, why ?}	\\\specialrule{0em}{1pt}{1pt}
		&$\texttt{pNet expression} ~~\lhook\joinrel\longrightarrow$ SMTLib expression	\\\specialrule{0em}{1pt}{1pt}
		&$assignments$  assert	\\\specialrule{0em}{1pt}{1pt}
		\specialrule{0em}{1pt}{1pt}\hline
	\end{tabular}
%\end{table}

@@

In the following definitions, we use abstract functions corresponding to
SMTlib constructs. In practice they can be implemented either as
SMTlib scripting programs, or as calls to the Z3 java API.

\subsubsection{Presentation Translation} %\TODO{Duplicate ? no,
                                %slightly more precise that in the
                                %paper body.}
We define here the translation of the algebra presentation, extended
with the constant operators collected from the pNet. It produces the
declaration of sorts (excepted Bool and Int) with their constructors
and selectors as \texttt{declare-datatypes}, and the declaration of
auxiliary operators as \texttt{declare-function} constructs.

For sorts, we must distinguish the case of mutually defined sorts
(e.g. edges and vertices in a graph), that must be declared within a
single \texttt{declare-datatypes} construct. For this we define a
dependency order on sort, and build the strongly connected components
of the graph of sorts:

\begin{definition}
Define the (strict) order "is-using" between sorts $S1~ is-using~ S2$ iff $S2$
occurs as the sort of one argument in the constructors of $S1$.
\end{definition}

\begin{lstlisting}
Let Pres = <Sorts, Constrs, Ops> an extended presentation
(i.e. including the constants from the pnet),
- define MySorts = Sorts \ {Bool, Int}
- compute the strongly connected components in the graph
  of MySorts with respect to the relation is-using
- for each SCC in this graph, construct the definitions, using
\texttt{declare-datatypes} for sorts, their constants and
constructors, and \texttt{declare-function} for the other operators,
using the following translation function:

\end{lstlisting}

@@

		& 			\\\specialrule{0em}{1pt}{1pt}
		$\dfrac{n = arity(constr) \neq 0 \qquad |selectors|=n
                  \qquad selectors =  BuildSels(constr)}
                       {constr \translate \texttt{(name($constr$) . $selectors$)}}$

@@

		& 			\\\specialrule{0em}{1pt}{1pt}
		$\dfrac{\EPres \vdash op : sort_1, ..., sort_n : sort_0
                \quad sortname_i = name(sort_i)}
                {op \translate \texttt{(declare-fun name($op$)
                    ($sortname_1 ... sortname_n$) $sortname_0$)}}$

@@

		& 			\\\specialrule{0em}{1pt}{1pt}
		$\dfrac{}
                       {FUN \translate \texttt{(map $\translate BuildFunInstance(CollectFunTypes(pNet))$) }}$

		%% &&       		\\\specialrule{0em}{1pt}{1pt}
		%% $\dfrac{}
                %%        {argstypes \translate
                %%          \texttt{(declare-fun $BuildFunInstance(argstypes)$}
                %%          \texttt{(map name $argstypes$) Action) }}$

@@

\smallskip
Where:
\begin{itemize}
  \item the $BuildSels$ function, for a constructor with $n$ arguments,
argument sorts $sort_i$, and selector names $sel_i$, builds the list
$\{(sel_i, sort_i)\}_{i\in[1..n]}$.
\item the $CollectFunTypes$ function collects all possible instances
  of the overloaded FUN argument types found in the pNet, as computed by the typing
  rules; and $BuildFunInstance$ use these argument types as suffixes
  to disambiguate the FUN operator name and build the corresponding
  \texttt{declare-fun} command.
\end{itemize}

\begin{example}
  For the CCS presentation above\TODO{Check}, we would get:
\begin{lstlisting}
  (declare-datatypes Data ())
  (declare-datatypes Channel ())
  (declare-datatypes ()
      ((Action (Emit (Chan_E Channel) (Value_E Data))
               (Receive (Chan_R Channel) (Value_R Data))
               Tau )))
  \end{lstlisting}
\end{example}

\begin{example}
  For our BIP use-case, we have built (see Fig.\ref{schema:smt-lib}):
  \begin{lstlisting}
  (declare-datatypes ()
      ((Action (fail) (resume) (timeout) (reset) (start) (tick) (ask)
    (FUN_Action_Bool (fst Action) (snd Bool))
    (Synchro (action Action)))))
  \end{lstlisting}
\end{example}

@@

\paragraph{properties}
The rules in the sections above are meant to guarantee that no runtime
errors would occur when running Z3 from our algorithm. More precisely,
we ensure that:
\begin{itemize}
\item all datatypes, functions, variable used are declared,
\item datatypes declared are well-founded,
\item all expressions in declarations and assertions are well-typed.
\end{itemize}

\section{Full output for the FailureTimer open automaton}
\label{appendix:fullOutput}
The following listing contains the final list of OTs generated for our
use-case. It is displayed in full and raw format, each OT containing
in its premisses: the list of transitions of pLTSs and actions of
holes, the predicate, the Post assignements. In the conclusion the
transition labelled by a fresh ``result'' variable.

\newpage
\begin{lstlisting}[basicstyle=\scriptsize\ttfamily, frame=single]
   {s0---failure(true)-->s1, t0---start(true)-->t1},{---:hb_B:13:1-->},   
   (failure(true)=failure(b1:sva_SV0:1:1))/\(start(true)=start(b2:sva_SV0:1:1))
   /\(:hb_B:13:1=fail(b0:sva_SV0:1:1))/\((b1:sva_SV0:1:1=b2:sva_SV0:1:1)
   /\(~(b1:sva_SV0:1:1\/b2:sva_SV0:1:1)\/b0:sva_SV0:1:1))
   /\(_fail_=:ra:1:1),{t := Max}
OT ---------------------------------------------------------------------------
        <s0_t0>-----:ra:1:1----><s1_t1>

@@

   {s1---resume(true)-->s0, t1---resume(true)-->t0},{---:hb_B:16:2-->},   
   (resume(true)=resume(b1:sva_SV1:1:2))/\(resume(true)=resume(b2:sva_SV1:1:2))
   /\(:hb_B:16:2=resume(b0:sva_SV1:1:2))/\((b1:sva_SV1:1:2=b2:sva_SV1:1:2)
   /\(~(b1:sva_SV1:1:2\/b2:sva_SV1:1:2)\/b0:sva_SV1:1:2))
   /\(_resume_=:ra:1:2),{}
OT ---------------------------------------------------------------------------
        <s1_t1>-----:ra:1:2----><s0_t0>

@@

\newpage
\begin{figure}[h]
$\begin{array}{r@{}l}
    ot_1  = &\openrule{\{s0 \xrightarrow{\OTvar{fail(true)}} s1, t0 \xrightarrow{\OTvar{start(true)}} t1\}, ~~\{\xrightarrow{hb_1}\},~~\OTvar{fail(false)}=\OTvar{fail(}\OTvar{b}_1\OTvar{)}\\
    \OTland \OTvar{start(true)}=\OTvar{start(}\OTvar{b}_2\OTvar{)}\OTland hb_1=\OTvar{fail(}\OTvar{b}_0\OTvar{)}\\
    \OTland v_1=\OTvar{\underline{fail}}\OTland \OTvar{b}_1 = \OTvar{b}_2
    \OTland \OTvar{(}\OTvar{b}_1 \lor \OTvar{b}_2\OTvar{)} \Rightarrow \OTvar{b}_0, ~~\{t := Max\}}
           {\ostate{s0,t0} \xrightarrow{v_1} \ostate{s1,t1}}\\
\end{array}$
  \smallskip

@@

\newpage
\subsection{Scaling up: several subsystems on the same bus}

\begin{figure}
    \centerline{\includegraphics[width=\linewidth]{XFIG/ScaleUp}}
    \caption{pNet encoding of the composition of two Failure Monitoring modules}  \label{schema:scaleup}
\end{figure}

\end{document}
